G\"odel's God in Isabelle/HOL

### 2013-10-29 Formalizations/Isabelle/GoedelGodSession/output/document/GoedelGod.tex
new Isabelle/HOL version of the proof

%
\begin{isabellebody}%
\def\isabellecontext{GoedelGod}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isamarkupsection{Introduction%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
A formalization and (partial) automation of Dana Scott's version \cite{ScottNotes}
 of Goedel's ontological argument \cite{GoedelNotes} in quantified modal logic KB (QML KB) is presented. 
 QML KB is in turn modeled as a fragment of classical higher-order logic (HOL). 
 Thus, the formalization is essentially a formalization in HOL. The employed embedding 
 of QML KB in HOL is adapting the work of Benzm\"uller and Paulson \cite{J23,B9}.
 Note that our QML KB formalization employs quantification over individuals and 
 quantification over sets of individuals (poperties).

 The formalization presented here has carried and formally verified in the Isabelle/HOL 
 proof assistant; for more information on this system we refer to the textbook by Nipkow, 
 Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
 at \url{http://isabelle.in.tum.de}.

 Some further notes:

@@

 code document at \url{??? todo ???} with the Isabelle build tool.
 \item The formalization presented here follows the THF \cite{J22} formalization available at: 
    \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/THF}
 \item The formalization is also related to the Coq \cite{Coq} formalization at:  
    \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Coq}

@@

    with Sledgehammer \cite{Sledgehammer} performing remote calls to Satallax 
    \cite{Satallax} and LEO-II \cite{LEO-II}. These calls then suggested respective 
    Metis \cite{Metis} calls as given below. Metis proofs are verified in Isabelle/HOL.
 \item The reconstruction with Metis still fails for Theorem T3, but when some further
    lemmata are introduced we get success; these lemmata are called help1-4.
 \item For consistency checking the Nitpick model finder \cite{Nitpick} has been employed.

@@

Types for possible worlds and for individuals.%

@@

An accessibility relation on worlds.%

@@

The B axiom (symmetry) for relation r; B is only needed for proving theorem T3.%

@@

 lifted to type $\sigma$. Further connectives could be introduced analogously.%

@@

\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ p\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline
\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymand}\ q\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymand}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymRightarrow}\ q\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymlongrightarrow}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline
\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}i\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}i\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}p{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}p\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ p\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ p\ v{\isacharparenright}{\isachardoublequoteclose}%

@@

For the grounding of lifted formulas the meta-predicate \isa{valid} is 
 introduced.%

@@

@@

The model finder Nitpick can now be used to confirm that the axioms and 
 definitions above are consistent. Unfortunately, the respective call-command for 
 Nitpick is not very intuitive.%
\isamarkupsection{The Formalization of G\"odel's Definitions and Axioms%
We introduce constant symbol \isa{P} (G\"odel's "Positive").%
The meaning of \isa{P} is constrained by axioms 
 \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$ (Either a property or 
 its negation is positive, but not both) and 
 \isa{A{\isadigit{2}}}: $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x) \imp \psi(x)]) 
 \imp P(\psi)]$ (A property necessarily implied by a positive property is positive).%
We prove theorem T1: $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ 
         (Positive properties are possibly exemplified).  
         T1 is proved directly by Sledghammer with command 
         \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}}. This successful 
         attempt then suggest to instead try the Metis call below.%
\ \ \ \ %
\ A{\isadigit{2}}\ A{\isadigit{1}}a\ \isacommand{unfolding}\isamarkupfalse%
\ mand{\isacharunderscore}def\ mbox{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ mnot{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
Next, we introduce the symbol \isa{G}, G\"odel's "God-like", which is defined 
We add axiom \isa{A{\isadigit{3}}}: $P(G)$ (The property of being God-like is positive) 
 and prove corollary \isa{C}: $\pos \ex x G(x)$ (Possibly, God exists). This
 this is directly proved by Sledgehammer, which then suggests the Metis call below.%
\ \ \ \ %
\ A{\isadigit{3}}\ T{\isadigit{1}}\ \isacommand{unfolding}\isamarkupfalse%
\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ valid{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
We introduce symbol \isa{ess}, for "Essence", which is defined as 
 and necessarily implying any of its properties).%
We now prove theorem \isa{T{\isadigit{2}}}: $\all x [G(x) \imp \ess{G}{x}]$ (Being 
 God-like is an essence of any God-like being). Again, Sledgehammer is used first, 
 which then suggests the Metis call below.%
\ \ \ \ %
\ A{\isadigit{1}}a\ A{\isadigit{1}}b\ A{\isadigit{4}}\ \isacommand{unfolding}\isamarkupfalse%
\ valid{\isacharunderscore}def\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mnot{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ mand{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ G{\isacharunderscore}def\ ess{\isacharunderscore}def\ \isanewline
\ \ \ \isacommand{by}\isamarkupfalse%
Now, symbol \isa{NE}, for "Necessary Existence", is introduced. It is 
 existence of an individual is the necessary exemplification of all its essences).%
We add axiom \isa{A{\isadigit{5}}}: $P(\NE)$ (Necessary existence is a positive property)%
The main theorem T3: $\nec \ex x G(x)$ (Necessarily, God exists) can be proven
 directly with Sledgehammer: 
 \isa{theorem\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequote}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequote}\ sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}} 
 , but the suggested Metis call then unfortunately still fails. In order to 
 obtain a full verification of the proof we therefore introduce some further lemmata.%
\ help{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ p{\isacharparenright}{\isacharbrackright}\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\isasymbox}\ p{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ %
\ sym\ \isacommand{unfolding}\isamarkupfalse%
\ valid{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ help{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ p{\isacharbrackright}\ {\isasymand}\ {\isacharbrackleft}p\ m{\isasymRightarrow}\ {\isasymbox}\ p{\isacharbrackright}\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ p{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ %
\isacommand{unfolding}\isamarkupfalse%
\ valid{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ \ \isanewline
\ help{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ {\isacharbrackleft}G\ x\ m{\isasymRightarrow}\ G\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ \ \ \ %
\ A{\isadigit{3}}\ A{\isadigit{5}}\ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}def\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ NE{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}metis\ {\isacharparenleft}lifting{\isacharcomma}\ mono{\isacharunderscore}tags{\isacharparenright}{\isacharparenright}%
\ help{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}i\ G\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ \ \ \ %
\ help{\isadigit{3}}\ T{\isadigit{2}}\ \isacommand{unfolding}\isamarkupfalse%
\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
Now, also Metis succeeds with proving theorem \isa{T{\isadigit{3}}}, and we also get
 \isa{T{\isadigit{4}}}:%
\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\ help{\isadigit{1}}\ help{\isadigit{2}}\ C\ help{\isadigit{4}}\ \isacommand{by}\isamarkupfalse%
Finally, we verify the consistency of our entire theory with Nitpick.%

@@

\paragraph{Acknowledgments:} Nik Sultana, Jasmin Blanchette and Larry Paulson provided very important 
 help wrt consistency checking in Isabelle. Jasmin Blanchette also showed us how to produce latex documents from Isabelle files.%

>>>>

%
\begin{isabellebody}%
\def\isabellecontext{GoedelGod}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isamarkupsection{Introduction%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
A formalization and (partial) automation of Dana Scott's version \cite{ScottNotes}
 of Goedel's ontological argument \cite{GoedelNotes} in quantified modal logic KB (QML KB) is presented. 
 QML KB is in turn modeled as a fragment of classical higher-order logic (HOL). 
 Thus, the formalization is essentially a formalization in HOL. The employed embedding 
 of QML KB in HOL is adapting the work of Benzm\"uller and Paulson \cite{J23,B9}.
 Note that the QML KB formalization employs quantification over individuals and 
 quantification over sets of individuals (poperties).

 The formalization presented here has been carried and formally verified in the Isabelle/HOL 
 proof assistant; for more information on this system see the textbook by Nipkow, 
 Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
 at the Isabelle homepage: \url{http://isabelle.in.tum.de}.

 Some further notes: \sloppy

@@

 code document at 
 \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle/GoedelGodSession} 
 with the Isabelle build tool.
 \item The formalization presented here is related to the THF \cite{J22} and 
    Coq \cite{Coq} formalizations available at
    \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/}.

@@

    with Sledgehammer \cite{Sledgehammer} performing remote calls to the higher-order automated
    theorem prover LEO-II \cite{LEO-II}. These calls then suggested respective 
    Metis \cite{Metis} calls as given below. The Metis proofs are then verified in Isabelle/HOL.
 \item For consistency checking, the Nitpick model finder \cite{Nitpick} has been employed.

@@

The types \isa{i} for possible worlds (or states) and \isa{mu} for individuals 
are introduced.%

@@

Possible worlds are connected by an accessibility relation .%

@@

The \isa{B} axiom (symmetry) for relation r is stated. \isa{B} is needed only 
for proving theorem T3.%

@@

lifted to type $\sigma$. Further connectives could be introduced analogously. \isa{definition} 
could be used instead of \isa{abbreviation}; the latter are always fully expanded/rewritten,
which is fine here, where the focus has been on proof automation, but which would lead to 
unnessarily complex proof tasks in a purely interactive session.%

@@

\ \ \isacommand{abbreviation}\isamarkupfalse%
\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ {\isasymPhi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline
\ \ \isacommand{abbreviation}\isamarkupfalse%
\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPhi}\ m{\isasymand}\ q\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymPhi}\ w\ {\isasymand}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
\ \ \isacommand{abbreviation}\isamarkupfalse%
\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymRightarrow}\ q\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymlongrightarrow}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline
\ \ \isacommand{abbreviation}\isamarkupfalse%
\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}i\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
\ \ \isacommand{abbreviation}\isamarkupfalse%
\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}i\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{abbreviation}\isamarkupfalse%
\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}p{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}p\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{abbreviation}\isamarkupfalse%
\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ p\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ p\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{abbreviation}\isamarkupfalse%
\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ p\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ p\ v{\isacharparenright}{\isachardoublequoteclose}%

@@

For the grounding of lifted formulas the meta-predicate \isa{valid} is introduced.%

@@

\ \ \isacommand{abbreviation}\isamarkupfalse%

@@

The model finder Nitpick confirms that the axioms and definitions above are consistent. 
Unfortunately, the respective command syntax for Nitpick is not very intuitive.%

@@

Constant symbol \isa{P} (G\"odel's "Positive") is introduced.%
The meaning of \isa{P} is restricted by axioms \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \phi 
[P(\neg \phi) \biimp \neg P(\phi)]$ (Either a property or its negation is positive, but not both.) 
and \isa{A{\isadigit{2}}}: $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x) \imp \psi(x)]) 
\imp P(\psi)]$ (A property necessarily implied by a positive property is positive.).%
We prove theorem T1: $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ (Positive 
properties are possibly exemplified). T1 is proved directly by Sledghammer with command \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}}. This successful attempt then suggest to 
instead try the Metis call in the line below. This Metis call generates a proof object that is 
verified in Isabelle/HOL's kernel.%
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
\ \ %
\ A{\isadigit{2}}\ A{\isadigit{1}}a\ \isacommand{by}\isamarkupfalse%
Next, the symbol \isa{G}, for "God-like", is introduced and defined 
Axiom \isa{A{\isadigit{3}}} is added: $P(G)$ (The property of being God-like is positive.).
Sledgehammer and Metis then prove corollary \isa{C}: $\pos \ex x G(x)$ 
(Possibly, God exists.).%
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
\ \ %
\ A{\isadigit{3}}\ T{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
Symbol \isa{ess}, for "Essence", is introduced and defined as 
and necessarily implying any of its properties.).%
Next, Sledgehammer and Metis prove theorem \isa{T{\isadigit{2}}}: $\all x [G(x) \imp \ess{G}{x}]$ 
(Being God-like is an essence of any God-like being).%
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
\ \ %
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}metis\ {\isacharparenleft}lifting{\isacharparenright}\ A{\isadigit{1}}b\ A{\isadigit{4}}\ G{\isacharunderscore}def\ ess{\isacharunderscore}def{\isacharparenright}%
Symbol \isa{NE}, for "Necessary Existence", is introduced and
existence of an individual is the necessary exemplification of all its essences.).%
Moreover, axiom \isa{A{\isadigit{5}}} is added: $P(\NE)$ (Necessary existence is a positive 
property.).%
Finally, Sledgehammer and Metis prove the main theorem \isa{T{\isadigit{3}}}: $\nec \ex x G(x)$ 
(Necessarily, God exists).%
\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
\ \ %
\ A{\isadigit{5}}\ C\ T{\isadigit{2}}\ sym\ G{\isacharunderscore}def\ NE{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ T{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}i\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
\ \ %
\ T{\isadigit{1}}\ T{\isadigit{3}}\ sym\ G{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
Finally, the consistency of our entire theory with Nitpick.%
\paragraph{Acknowledgments:} Nik Sultana, Jasmin Blanchette and Larry Paulson provided 
very important help wrt consistency checking in Isabelle. Jasmin Blanchette instructed us on how to 
produce latex documents from Isabelle sources, and he pointed to many useful tricks in Isabelle.%
